Prizm Silver Rookie Cards

### 2024-04-25 paper/paper.qmd
touching up

---
title: "Performance of Rookie Can Predict Price of Their Rookie Prizm Silver Cards"
subtitle: "My subtitle if needed"
author: 
  - Harrison Huang
thanks: "Code and data are available at: https://github.com/lemonface88/NBA"
date: today 
date-format: long
abstract: "Collecting and investing in cards has been relevant for a while now. There are complex factors that contribute to the price fluxation of these cards. This analysis will use player statistics to create a model to try to grasp possible future price trends of the card. A linear model was used in this analysis. The results indicated that player statistics may still not be sufficient to predict future prices of the rookie card."
format: pdf
number-sections: true
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false

@@

library(readr)
#write_bib()

@@

The first place the data was collected from is ESPN, player statistics were obtained from scrapping the website and picking out the right tables. Please be respectful if you decide to use it for any other purpose. Each player had multiple tables and I downloaded the tables and merged players’ 202202-2023 and 2023-2024 data. This allows all statistics to be viewed in one table. In addition, when considering which of the statistics to focus on, I asked an NBA reporter Lisa Kao for suggestions. She mentioned that average points, blocks, steals, and assists are the top 5 statistics focused on in the industry. Something to keep in mind is that these statistics presented are all positive statistics, which means that the higher the number the better the player has performed.

@@

Below is a brief explination of the tables including player statistics:

@@

kable(anthony_edward, padding = 5)
```

@@

## Price of increasing performance

@@

## Third discussion point

@@

Since there isn’t a major platform that provides the tracking of every transaction of a card, it is very difficult to obtain the true price of the card. Sportscardpro only gathers its data from completed transactions of that card on eBay, which is only a portion of the sales that happen on the internet. Transactions of the card that often happen offline in stores are also not tracked, if I sold a card to my friend for a cheaper price, it is not reflected in the dataset from Sportscardpro.

@@

```{r}
#| include: true
#| eval: true
#| echo: false
#| message: false
#| warning: false

@@

posterior_vs_prior(maxey_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()

@@

posterior_vs_prior(haliburton_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()

>>>>

---
title: "Prizm Silver Rookie Cards"
subtitle: "An Analysis using statistic performance of a specific player and their rookie card price"
author: 
  - Harrison Huang
thanks: "Code and data are available at: https://github.com/lemonface88/NBA"
date: today 
date-format: long
abstract: "Collecting and investing in cards has been relevant for a while now. There are complex factors that contribute to the price fluxation of these cards. This analysis will use player statistics to create a model to try to grasp possible future price trends of the card. A linear model was used in this analysis. The results indicated that player statistics may still not be sufficient to predict future prices of the rookie card."
format: pdf
number-sections: true
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false

@@

library(readr)

@@

The first place the data was collected from is @ESPN, player statistics were obtained from scrapping the website and picking out the right tables. Please be respectful if you decide to use it for any other purpose. Each player had multiple tables and I downloaded the tables and merged players’ 202202-2023 and 2023-2024 data. This allows all statistics to be viewed in one table. In addition, when considering which of the statistics to focus on, I asked an NBA reporter Lisa Kao for suggestions. She mentioned that average points, blocks, steals, and assists are the top 5 statistics focused on in the industry. Something to keep in mind is that these statistics presented are all positive statistics, which means that the higher the number the better the player has performed.

@@

Below is a brief explanation of the tables including player statistics:

@@

kable(anthony_edward, padding = 5)

@@

## Tyrese Maxey Preduction
```{r}
#| label: fig-1
#| fig-cap: The comparision between past season versus predicted improving season of Tyrese Maxey
#| fig-subcap: ["Model Predicted Price", "Price from 2022-2024 season"]
#| echo: false
#| warning: false
#| layout-ncol: 2

maxey_prediction_table |> ggplot(aes(x = `Month`, y = `Predicted Price`)) + 
  geom_line(group =1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Predicted price of future seasons with improving performance",
    x = "Dates",
    y = "Price"

  )

price_data |> ggplot(aes(x = `Date`, y = `Tyrese Maxey`)) + 
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Price from 2022-2024 season",
    x = "Dates",
    y = "Price"

  )

## Tyrese Haliburton Preduction
```{r}
#| label: fig-2
#| fig-cap: The comparision between past season versus predicted improving season of Tyrese Haliburton
#| fig-subcap: ["Model Predicted Price", "Price from 2022-2024 season"]
#| echo: false
#| warning: false
#| layout-ncol: 2

library(ggplot2)

haliburton_prediction_table |> ggplot(aes(x = `Month`, y = `Predicted Price`)) + 
  geom_line(group =1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Predicted price of future seasons with improving performance",
    x = "Dates",
    y = "Price"

  )+ 
  geom_point() +
  geom_smooth(method=lm)

price_data |> ggplot(aes(x = `Date`, y = `Tyrese Haliburton`)) + 
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Price from 2022-2024 season",
    x = "Dates",
    y = "Price"

  )

## Anthony Edward Prediction

```{r}
#| label: fig-3
#| fig-cap: The comparision between past season versus predicted improving season of Anthony Edward
#| fig-subcap: ["Model Predicted Price", "Price from 2022-2024 season"]
#| echo: false
#| warning: false
#| layout-ncol: 2

anthony_prediction_table |> ggplot(aes(x = `Month`, y = `Predicted Price`)) + 
  geom_line(group =1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Predicted price of future seasons with improving performance",
    x = "Dates",
    y = "Price"

  )

price_data |> ggplot(aes(x = `Date`, y = `Anthony Edward`)) + 
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Price from 2022-2024 season",
    x = "Dates",
    y = "Price"

  )

In @fig-1, @fig-2, and @fig-3 the results are all very similar. The over all price has a downward trend initially without the use of the model. However, in the start of the 2023 season, all three players card price saw an upward trend with then later died down again falling into another downward trend. This could be because if people coming back to the sport also bringing more attention to playing cards ultimately leading to increase in the price. When the month jumps from April to October, a price increase can be observed, which indicates an end of a season to the start of the next season. The predicted price interestingly displayed an increase at the beginning of each season and decreasing of price after around December of each season.

@@

## Season timing
After taking a closer look at trends from both predicted and past data, an interesting point came up mind. When observing turning points of the price, they all seems to be the start and end of a season. Price increasing trends tends to happen at the start of each season, and then become more and more unpredictable throughout the season. This brings attention to the timing of the season and the price may be quite related and even more than the player statistics. Because despite implementing an increasing performance for each of the player statistics to run the model on, there are no clear signs of increasing price overall. The thing that stood out the most is the price increase at the start of every season in October. This could suggest that the initial hype of a new season could be contrbuting to the incresae of the card price.

@@

Since there isn’t a major platform that provides the tracking of every transaction of a card, it is very difficult to obtain the true price of the card. Sportscardpro only gathers its data from completed transactions of that card on eBay, which is only a portion of the sales that happen on the internet. Transactions of the card that often happen offline in stores are also not tracked, if I sold a card to my friend for a cheaper price, it is not reflected in the data set from @Sportscardpro.

@@

```{r}
#| label: fig-4
#| fig-cap: Model fit of Tyrese Maxey
#| echo: false
#| warning: false
#| layout-ncol: 2

@@

posterior_vs_prior(maxey_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
```{r}
#| label: fig-5
#| fig-cap: Model fit of Tyrese Haliburton
#| echo: false
#| warning: false
#| layout-ncol: 2

@@

posterior_vs_prior(haliburton_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| label: fig-6
#| fig-cap: Model fit of Anthony Edward
#| echo: false
#| warning: false
#| layout-ncol: 2
